How has the Brazilian Amazon been constructed as a problem? An analysis of presidential speeches since 1985

### 2022-04-05 Draft_April_2022/Full_draft_20220401.tex
#preliminary final draft

---
title: How has the Brazilian Amazon been constructed as a problem? An analysis of
  presidential speeches since 1985
author:
- Livio Silva-Muller^[Phd Candidate, The Geneva Graduate Institute , livio.silva@graduateinstitute.ch]
- Henrique Sposito^[Phd Candidate, The Geneva Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
date: "April 2022"
output:
  word_document:
    reference_docx: template.docx
  pdf_document: default
  html_document:
    df_print: paged
bibliography: references.bib
csl: chicago.csl
abstract: This paper ...
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(skimr)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(nnet)
library(stargazer)

Sys.setlocale("LC_ALL", "C")

options(stringsAsFactors = FALSE, scipen = 999)

```

```{r data, message=FALSE, warning=FALSE, include=FALSE}

pop<- readRDS("~/Documents/GitHub/amazondef/BR_presid_speeches_final.Rds")
ama <- readRDS("~/Documents/GitHub/amazondef/final_data_as.Rds")

ama <-  ama %>%
          select(-c(title, date, text, false_positives)) %>%
          rename (ei= "EI") %>%
          rename (sd= "SD")

pop <-  pop %>%
          select(-c(title, date, text, location))

```

```{r data prep for prob_construction, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

sov <- ama %>% 
  filter(sov==1) %>% 
  count(sov, year) %>%
  group_by(year) %>%
  select(-c(sov))%>%
  rename (sov = n)

ei <- ama %>% 
  filter(ei==1) %>% 
  count(ei, year) %>%
  group_by(year) %>%
  select(-c(ei))%>%
  rename (ei = n) %>%
  ungroup() %>%
  tibble::add_row(year = 2017, ei = 0) %>%
  tibble::add_row(year = 2018, ei = 0)

sd <- ama %>% 
  filter(sd==1) %>% 
  count(sd, year) %>%
  group_by(year) %>%
  select(-c(sd))%>%
  rename (sd = n)

con <- ama %>% 
  filter(con==1) %>% 
  count(con, year) %>%
  group_by(year) %>%
  select(-c(con))%>%
  rename (con = n)

other <- ama %>% 
  filter(other==1) %>% 
  count(other, year) %>%
  group_by(year) %>%
  select(-c(other))%>%
  rename (other = n)

temp <- ei %>% left_join(other) %>% 
                left_join(con) %>%
                left_join(sd) %>%
                left_join(sov) %>%
                replace(is.na(.), 0)

```

```{r data prep for mixed_types, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

ama_mx <- ama %>% 
  unite(mixed_type, sov:con, sep="_") %>%
  mutate(mx_cat= case_when (mixed_type=='1_0_0_0' ~ "Pure Sovereignty",
                            mixed_type=='0_1_0_0' ~ "Pure Economic Integration",
                            mixed_type=='0_0_1_0' ~ "Pure Social Development",
                            mixed_type=='0_0_0_1' ~ "Pure Conservation",
                            mixed_type=='0_0_0_0' ~"Other",
                            mixed_type=='0_1_0_1' ~"Economic Conservation",
                            mixed_type=='0_1_1_0' ~"Economic and Social Development",
                            mixed_type=='1_0_0_1' ~"Sovereign Conservation",
                            mixed_type=='0_1_1_1' ~ "Sustainable Development",
                            mixed_type=='1_1_0_0'~ "Sovereign Integration",
                            mixed_type=='0_0_1_1' ~"Social Conservation",
                            mixed_type=='1_1_0_1' ~"Other Mixed-types",
                            mixed_type=='1_0_1_0' ~"Other Mixed-types",
                            mixed_type=='1_0_1_0' ~"Other Mixed-types",
                            mixed_type=='1_1_1_1' ~"Other Mixed-types",
                            mixed_type=='1_1_1_0' ~"Other Mixed-types",
                            mixed_type=='1_0_1_1' ~"Other Mixed-types"))%>%
  mutate(mx_cat=factor(mx_cat, levels = c("Pure Economic Integration", "Pure Conservation", "Pure Social Development", "Pure Sovereignty", "Other", "Economic Conservation", "Economic and Social Development", "Sovereign Conservation","Sustainable Development","Sovereign Integration", "Social Conservation", "Other Mixed-types"))) %>%
mutate(mx_cat2= case_when (mixed_type=='1_0_0_0' ~ "Pure-type",
                            mixed_type=='0_1_0_0' ~ "Pure-type",
                            mixed_type=='0_0_1_0' ~ "Pure-type",
                            mixed_type=='0_0_0_1' ~ "Pure-type",
                            mixed_type=='0_0_0_0' ~"Other",
                            mixed_type=='0_1_0_1' ~"Mixed-type",
                            mixed_type=='0_1_1_0' ~"Mixed-type",
                            mixed_type=='1_0_0_1' ~"Mixed-type",
                            mixed_type=='0_1_1_1' ~ "Mixed-type",
                            mixed_type=='1_1_0_0'~ "Mixed-type",
                            mixed_type=='0_0_1_1' ~"Mixed-type",
                            mixed_type=='1_1_0_1' ~"Mixed-type",
                            mixed_type=='1_0_1_0' ~"Mixed-type",
                            mixed_type=='1_0_1_0' ~"Mixed-type",
                            mixed_type=='1_1_1_1' ~"Mixed-type",
                            mixed_type=='1_1_1_0' ~"Mixed-type",
                            mixed_type=='1_0_1_1' ~"Mixed-type"))

```

@@

# 1 Introduction

The Amazon needs to be protected from foreign interests. The Amazon needs to be exploited for its natural resources. The Amazon should be preserved as a standing ecosystem. Historically, different Brazilian federal government proposed diverse policies to deal with the Amazon. Each of these policies contain an implicit assumption of what needs to be solved, or in other words, it represents the region, the forest, or the peoples as a particular problem. In the three examples above, the Amazon is represented as an issue of national sovereignty, economic integration, and environmental conservation, respectively. Each of these constructions, and their proposed solutions, have been described as policy cycles of Brazilian governments [@acker2014; @hecht1990; @hochstetler2007]. However, policy cycles are usually represented monolithically, advancing a view that specific governments see the Amazon as an instance of only one specific problem. Albeit the current academic calls to understand the environment as a social-cultural construction and to identify the effect of culture for environmental outcomes [@lepolaindewaroux2021], we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem over time and across locations, and how this varies.

In this article, we investigate how the Brazilian Amazon has been constructed as a problem in political discourses. Building on @hirschman1963 concept of chosen problems in policy making, we propose a framework to identify how problem-construction varies over time, by location, and between and within governments. Although problem-construction takes place in a series of instances (e.g. policy committees, legislative bodies, media, etc.), we analyze the case of political discourse by Brazilian presidents since 1985. We opt for presidential speeches for three reasons. First, political discourses at the top have the power to introduce and justify public policy, as well as shape its perception to broad audiences [@zarefsky2004]. In turn, policy perception is key for policy adoption and implementation [@alesina2009]. The literature has shown that deforestation rates in Brazil are more responsive to the government's environmental policy than exogenous factors as market fluctuations [@assunção2015; @capobianco2019; @capobianco2021]. Thus, understanding how policy comes about discursively is important. Second, environmental discourse at the top can help expand or restrict what types of behaviors are accepted in the ground. When Brazilian presidents speak about the Amazon it not only makes headlines, nationally and internationally, but also incites responses, shapes expectations, and feeds into the behavior of many actors involved with the Amazon, from investors to agribusiness to local farmer[@brice2021; @harris2021; @miranda2021]. This is especially pertinent for deforestation as previous research found that policy expectations, generated from material and discursive governmental practices, are a crucial factor in decisions to deforest at the ground [@capobianco2019; @campbell2015]. Unpacking discourse at the top might help us raise hypotheses about environmental outcomes that are culturally situated. Finally, as our theoretical framework suggests, problem-construction varies by geographic location. Presidential discourses take place in a series of sites with diverse audiences: from launching a new bridge in a small municipality in the middle of the Amazon, to a keynote speech in a business association in São Paulo, to the UN general assembly in New York city. Working with presidential discourses allows us to identify this variation in meaningful ways and better how the Amazon is socially constructed.

@@

# 2 Amazonian policy-cycles, discourse, and problem-construction

## 2.1 Literature Review: policy-cycles in the Amazonian literature

@@

### 2.1.1 National sovereignty

In The Fate of the Forest: Developers, Destroyers, and Defenders of the Amazon, Hecht and Cockburn write that all over the world tropical forests are destroyed, but "what imbues the case of the Amazon with such passion is the symbolic content of the dreams it ignites" [-@hecht1990, pp.1]. It started with the first natural history of the New World, by Oviedo in 1535, who recounts the stories of conquest of local populations and gold hoarders. The dream of fortunes to be found in the Eldorado composed the imaginaries of *bandeirantes*[^1] from the southeast of Brazil and colonizers from everywhere else. It rendered the territory the venue for aspiration and object of an intense scramble in the subsequent centuries, defined as "a (...) form of nation building (...)" (Hecht and Cockburn 1990, preface). The Portuguese empire and subsequently the Brazilian monarchy were concerned with establishing their territory. In the process of securing Amazonian borders, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht2013, pp.8], and when the dust settled and the scramble was over, half of the Amazon emerged Brazilian. While Brazilian military diplomacy was very successful, the process did not come without its traumas. A significant experience were the negotiations with Bolivia in 1902 to secure the Amazonian state of Acre, during which they found out about American attempts to trick Brazil[@hecht1990]. This case was still part of the memory of the generals who led the country during the military dictatorship of 1964 and wanted to protect Brazil's sovereignty over the Amazon from the communist threat during the Cold War[@garfield2013]. As we move from a world where non-state actors gain importance in environmental governance and international politics generally [@silva-muller2022; @andonova2014; @westerwinter2021], the sovereignty problem becomes more varied. Multiple non-state actors (NGOs, foundations, IOs, and so on) join the conversation about Amazonian policies more substantially as the military dictatorship starts to end [@hochstetler2021; @capobianco2019; @franchini2019]. Threads to national sovereignty, consequently, can be interpreted as coming from a different set of actors than before. False claims about Brazilian policies in international and domestic fora, for instance, are often tied to strategies of 'internationalizing' the Amazon. Relatedly, mentions of Amazonian myths which have been debunked as the 'Earth of the Lungs', are also tied to internationalizing strategies.

[^1]: B*andeirantes* means "flag-carriers", the word is used to designate Portuguese colonials and later Brazilian explorers, expanding the Brazilian territory beyond what the Tordesillas Treaty established. The treaty allocated almost the whole Amazonian territory to the Spanish Empire, the bandeirantes took much of the territory afterwards.

@@

### 2.1.2 Economic integration

The dictatorships of Vargas (1937-46) and the military (1964-89) took over the task of modernizing the Amazon. In 1966, the Brazilian Military dictatorship launched Operation Amazon, a policy to modernize the region based on a set of assumptions [@acker2014]. First, nature should be conquered by men. Second, exploiting natural resources would render the Amazon region a global powerhouse. Third, such a project would integrate the region with the rest of the country. Concretely, this meant a series of infrastructure projects, such as roads and dams, incentives for settlers to develop ranches and expand the agricultural frontier, as well as establishing tax free zones to attract industry. The capital to conduct such changes, paradoxically, came from national and international sources [@acker2014], leading to a series of national and international enterprises settling in the Amazon region. @capobianco2019 describes the period from the 1950-80 in a similar fashion, referring to a wider range of policies of economic integration: the 1953 Plano de Valorização Econômica da Amazônia; the 1966 Superintendência do Desenvolvimento da Amazônia; the 1967 Superintendência da Zona Franca de Manaus; the 1970 Plano de Integração Nacional; the 1975 Programa Polamazônia; the 1980s Programa Grande Carajás and Programa Calha Norte; among others.

The economic integration problem advances the view that the Brazilian Amazon needs to be developed and modernized. The policy solution relates to the creation of a series of policies, often centralized by the federal government and thus external to the region [@becker2005], that have at its core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region in the national and international economy.

### 2.1.3 Environmental conservation

The rapid economic changes in the region in the 1960s, 70s, and 80s were matched with the birth of environmental institutions such as the New Forest Code (1964), the Secretary of Environment (1973), and the National Environment Law (1980) [@drummond2006]. A common explanation for these institutions in the Amazonian literature is the impression of lack of control after years of economic integration. [@acker2021; @capobianco2021; @hecht1990]. As deforestation, fires, and violence rose in the region, catching international attention, the military government deemed as necessary the establishment of an environmental bureaucracy. This process accelerated in the late 1980s, with the birth of modern environmentalism epitomized in the 1992 Earth Summit in Rio de Janeiro (Hochstetler 2021; Capobianco, 2019; Hochstetler and Keck, 2007). @hochstetler2007 argue that during preparations for the summit, a new form of Brazilian environmentalism emerged: socio-environmentalism. They define it as an emphasis on local livelihoods of people while protecting nature. @capobianco2019 argues in a similar line, establishing socio-environmentalism as the main government response in the 1990s and early 2000s in a series of policies: the 2001 Sistema Nacional de Unidades de Conservação; the 2003 Programa Amanônia Sustentavel; the 2004 Plano de Ação para a Prevenção e Controle do Desmatamento na Amazônia Legal; the 2004 Plano BR-163 Sustentável; the 2010 Lei Nacional das Mudanças Climáticas; among others.

The conservationist problem-construction advances the view that Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies more investment in command-and-control infrastructure (as remote-sensing technology for environmental outcome measurement), more investment in the valuation of standing eco-systems through incentive schemes, and more policies facilitating indigenous environmental practices.

### 2.1.4 Policy and discourse

@@

Problem-construction at the level of discourse is also more varied. They are not monolithic in time, across location, or even by the same speaker. While governmental discourses in Brazil have been studied for topic such as inflation or race relations , we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that in congress the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

While we find no major studies systematically looking into Amazonian related themes in Brazilian presidential speeches, we find, for example, several analyzes of environmental discourse in American presidential speeches. @calderwood2019 examines 2919 mentions of climate change in American official presidential speeches since 1989. Among various findings, one that stands out is that American presidents frequently side-step the environmental aspects of climate change (ibid). He also identifies a shift from economic to security framing of climate issues, side-lining its environmental aspects. In another article, @calderwood2020 tests the effect of geographic location and type of communication regarding climate change. Building prominently on @putnam1988 but also others, he hypothesizes that presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. He finds evidence in support of his hypothesis, suggesting American presidential discourse at the top on climate change does change based on location. Another example is @bevitori2015, who investigates how the 'environment' has been constructed in American presidential discourse since 1960 using a more automated approach. The author finds that mentions of the environment are typically co-selected with the pronoun 'our', as well as with 'economy', 'clean', and 'preserve'.

@@

## 2.2 Theoretical framework: problem-construction and presidential discourse

In "Journey towards Progress", Hirschman (1963) analyzes three policy problems in three different Latin American countries. The author draws a conceptual distinction between pressing problems (pressured from outside parties to the government) and chosen problems (chosen by the government at their own discretion). Pressing problems can be either privileged or neglected depending on the degree of pressure exercised by the interested group. Problems can change from pressing to chosen across time and in space as a function of solutions becoming available, changing level of government control in society, or top policymakers shifting interests[@hirschman1963, pp.388-91]. Choosing a problem, though, entails a decision on how to represent it[@bacchi2009]. As Bacchi puts "policy has a cultural dimension. It takes shape within specific historical and national or international contexts" [-@bacchi2009, pp.10]. The existence or proposal of a policy generally implies that there is a (public) problem that needs (governmental) action to be fixed (ibid). The alleged problem is not always explicitly stated in policy. Hirschman exemplifies chosen problems with the case of the construction of Brasilia [-@hirschman1975, pp.388]. But building Brasilia can solve a problem of regional inequality, a problem of a dormant economy without state investment, a problem of political representation, or all three. How to represent a policy is a matter of choice. And different representations speak to different audiences.

@@

# 3 Research Design

## 3.1 Data and modeling: operationalizing Amazonian problem construction

## 3.2 Analysis and limitations

# 4 How has the amazon been constructed as a problem?

@@

## 4.1 The rises and falls of the Amazon as a topic in presidential speeches

@@

## 4.2 Amazonian problem-construction in time

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={How has the Brazilian Amazon been constructed as a problem? An analysis of presidential speeches since 1985},
  pdfauthor={Livio Silva-Muller; Henrique Sposito},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{How has the Brazilian Amazon been constructed as a problem? An
analysis of presidential speeches since 1985}
\author{Livio Silva-Muller\footnote{Phd Candidate, The Graduate
  Institute Geneva,
  \href{mailto:livio.silva@graduateinstitute.ch}{\nolinkurl{livio.silva@graduateinstitute.ch}}} \and Henrique
Sposito\footnote{Phd Candidate, The Graduate Institute Geneva,
  \href{mailto:henrique.sposito@graduateinstitute.ch}{\nolinkurl{henrique.sposito@graduateinstitute.ch}}}}
\date{April 2022}

\begin{document}
\maketitle
\begin{abstract}
This paper \ldots{}
\end{abstract}

@@

\hypertarget{introduction}{%
\section{1 Introduction}\label{introduction}}

The Amazon needs to be protected from foreign interests. The Amazon
needs to be exploited for its natural resources. The Amazon should be
preserved as a standing ecosystem. Historically, different Brazilian
federal government proposed diverse policies to deal with the Amazon.
Each of these policies contain an implicit assumption of what needs to
be solved, or in other words, it represents the region, the forest, or
the peoples as a particular problem. In the three examples above, the
Amazon is represented as an issue of national sovereignty, economic
integration, and environmental conservation, respectively. Each of these
constructions, and their proposed solutions, have been described as
policy cycles of Brazilian governments (Acker 2014; Hecht and Cockburn
1990; Hochstetler and Keck 2007). However, policy cycles are usually
represented monolithically, advancing a view that specific governments
see the Amazon as an instance of only one specific problem. Albeit the
current academic calls to understand the environment as a
social-cultural construction and to identify the effect of culture for
environmental outcomes (Polain de Waroux et al. 2021), we lack empirical
accounts of how the Brazilian Amazon has been constructed as a problem
over time and across locations, and how this varies.

In this article, we investigate how the Brazilian Amazon has been
constructed as a problem in political discourses. Building on Hirschman
(1963) concept of chosen problems in policy making, we propose a
framework to identify how problem-construction varies over time, by
location, and between and within governments. Although
problem-construction takes place in a series of instances (e.g.~policy
committees, legislative bodies, media, etc.), we analyze the case of
political discourse by Brazilian presidents since 1985. We opt for
presidential speeches for three reasons. First, political discourses at
the top have the power to introduce and justify public policy, as well
as shape its perception to broad audiences (Zarefsky 2004). In turn,
policy perception is key for policy adoption and implementation (Alesina
and Giuliano 2009). The literature has shown that deforestation rates in
Brazil are more responsive to the government's environmental policy than
exogenous factors as market fluctuations (Assunção, Gandour, and Rocha
2015; Capobianco 2019, 2021). Thus, understanding how policy comes about
discursively is important. Second, environmental discourse at the top
can help expand or restrict what types of behaviors are accepted in the
ground. When Brazilian presidents speak about the Amazon it not only
makes headlines, nationally and internationally, but also incites
responses, shapes expectations, and feeds into the behavior of many
actors involved with the Amazon, from investors to agribusiness to local
farmer(Brice and Smith 2021; Harris 2021; Miranda 2021). This is
especially pertinent for deforestation as previous research found that
policy expectations, generated from material and discursive governmental
practices, are a crucial factor in decisions to deforest at the ground
(Capobianco 2019; Campbell 2015). Unpacking discourse at the top might
help us raise hypotheses about environmental outcomes that are
culturally situated. Finally, as our theoretical framework suggests,
problem-construction varies by geographic location. Presidential
discourses take place in a series of sites with diverse audiences: from
launching a new bridge in a small municipality in the middle of the
Amazon, to a keynote speech in a business association in São Paulo, to
the UN general assembly in New York city. Working with presidential
discourses allows us to identify this variation in meaningful ways and
better how the Amazon is socially constructed.

@@

\hypertarget{amazonian-policy-cycles-discourse-and-problem-construction}{%
\section{2 Amazonian policy-cycles, discourse, and
problem-construction}\label{amazonian-policy-cycles-discourse-and-problem-construction}}

\hypertarget{literature-review-policy-cycles-in-the-amazonian-literature}{%
\subsection{2.1 Literature Review: policy-cycles in the Amazonian
literature}\label{literature-review-policy-cycles-in-the-amazonian-literature}}

@@

\hypertarget{national-sovereignty}{%
\subsubsection{2.1.1 National sovereignty}\label{national-sovereignty}}

In The Fate of the Forest: Developers, Destroyers, and Defenders of the
Amazon, Hecht and Cockburn write that all over the world tropical
forests are destroyed, but ``what imbues the case of the Amazon with
such passion is the symbolic content of the dreams it ignites'' (1990,
1). It started with the first natural history of the New World, by
Oviedo in 1535, who recounts the stories of conquest of local
populations and gold hoarders. The dream of fortunes to be found in the
Eldorado composed the imaginaries of \emph{bandeirantes}\footnote{B\emph{andeirantes}
  means ``flag-carriers'', the word is used to designate Portuguese
  colonials and later Brazilian explorers, expanding the Brazilian
  territory beyond what the Tordesillas Treaty established. The treaty
  allocated almost the whole Amazonian territory to the Spanish Empire,
  the bandeirantes took much of the territory afterwards.} from the
southeast of Brazil and colonizers from everywhere else. It rendered the
territory the venue for aspiration and object of an intense scramble in
the subsequent centuries, defined as ``a (\ldots) form of nation
building (\ldots)'' (Hecht and Cockburn 1990, preface). The Portuguese
empire and subsequently the Brazilian monarchy were concerned with
establishing their territory. In the process of securing Amazonian
borders, Brazil thwarted ``the imperial ambitions of France, Britain,
the United States, Belgium, Bolivia, and Peru'' (Hecht 2013, 8), and
when the dust settled and the scramble was over, half of the Amazon
emerged Brazilian. While Brazilian military diplomacy was very
successful, the process did not come without its traumas. A significant
experience were the negotiations with Bolivia in 1902 to secure the
Amazonian state of Acre, during which they found out about American
attempts to trick Brazil(Hecht and Cockburn 1990). This case was still
part of the memory of the generals who led the country during the
military dictatorship of 1964 and wanted to protect Brazil's sovereignty
over the Amazon from the communist threat during the Cold War(Garfield
2013). As we move from a world where non-state actors gain importance in
environmental governance and international politics generally
(Silva-Muller and Faul 2022; Andonova 2014; Westerwinter 2021), the
sovereignty problem becomes more varied. Multiple non-state actors
(NGOs, foundations, IOs, and so on) join the conversation about
Amazonian policies more substantially as the military dictatorship
starts to end (Hochstetler 2021; Capobianco 2019; Franchini and Viola
2019). Threads to national sovereignty, consequently, can be interpreted
as coming from a different set of actors than before. False claims about
Brazilian policies in international and domestic fora, for instance, are
often tied to strategies of `internationalizing' the Amazon. Relatedly,
mentions of Amazonian myths which have been debunked as the `Earth of
the Lungs', are also tied to internationalizing strategies.

@@

\hypertarget{economic-integration}{%
\subsubsection{2.1.2 Economic integration}\label{economic-integration}}

The dictatorships of Vargas (1937-46) and the military (1964-89) took
over the task of modernizing the Amazon. In 1966, the Brazilian Military
dictatorship launched Operation Amazon, a policy to modernize the region
based on a set of assumptions (Acker 2014). First, nature should be
conquered by men. Second, exploiting natural resources would render the
Amazon region a global powerhouse. Third, such a project would integrate
the region with the rest of the country. Concretely, this meant a series
of infrastructure projects, such as roads and dams, incentives for
settlers to develop ranches and expand the agricultural frontier, as
well as establishing tax free zones to attract industry. The capital to
conduct such changes, paradoxically, came from national and
international sources (Acker 2014), leading to a series of national and
international enterprises settling in the Amazon region. Capobianco
(2019) describes the period from the 1950-80 in a similar fashion,
referring to a wider range of policies of economic integration: the 1953
Plano de Valorização Econômica da Amazônia; the 1966 Superintendência do
Desenvolvimento da Amazônia; the 1967 Superintendência da Zona Franca de
Manaus; the 1970 Plano de Integração Nacional; the 1975 Programa
Polamazônia; the 1980s Programa Grande Carajás and Programa Calha Norte;
among others.

The economic integration problem advances the view that the Brazilian
Amazon needs to be developed and modernized. The policy solution relates
to the creation of a series of policies, often centralized by the
federal government and thus external to the region (Becker 2005), that
have at its core the development of the necessary infrastructure
(physical, fiscal, or monetary) to integrate the region in the national
and international economy.

\hypertarget{environmental-conservation}{%
\subsubsection{2.1.3 Environmental
conservation}\label{environmental-conservation}}

The rapid economic changes in the region in the 1960s, 70s, and 80s were
matched with the birth of environmental institutions such as the New
Forest Code (1964), the Secretary of Environment (1973), and the
National Environment Law (1980) (Drummond and Barros-Platiau 2006). A
common explanation for these institutions in the Amazonian literature is
the impression of lack of control after years of economic integration.
(Acker 2021; Capobianco 2021; Hecht and Cockburn 1990). As
deforestation, fires, and violence rose in the region, catching
international attention, the military government deemed as necessary the
establishment of an environmental bureaucracy. This process accelerated
in the late 1980s, with the birth of modern environmentalism epitomized
in the 1992 Earth Summit in Rio de Janeiro (Hochstetler 2021;
Capobianco, 2019; Hochstetler and Keck, 2007). Hochstetler and Keck
(2007) argue that during preparations for the summit, a new form of
Brazilian environmentalism emerged: socio-environmentalism. They define
it as an emphasis on local livelihoods of people while protecting
nature. Capobianco (2019) argues in a similar line, establishing
socio-environmentalism as the main government response in the 1990s and
early 2000s in a series of policies: the 2001 Sistema Nacional de
Unidades de Conservação; the 2003 Programa Amanônia Sustentavel; the
2004 Plano de Ação para a Prevenção e Controle do Desmatamento na
Amazônia Legal; the 2004 Plano BR-163 Sustentável; the 2010 Lei Nacional
das Mudanças Climáticas; among others.

The conservationist problem-construction advances the view that Amazon
should be preserved, deforestation should be halted, and the sustainable
practices of indigenous and local peoples should be maintained through
protection of their territories and rights to self-determination
(Hochstetler and Keck 2007). The policy solution implies more investment
in command-and-control infrastructure (as remote-sensing technology for
environmental outcome measurement), more investment in the valuation of
standing eco-systems through incentive schemes, and more policies
facilitating indigenous environmental practices.

\hypertarget{policy-and-discourse}{%
\subsubsection{2.1.4 Policy and discourse}\label{policy-and-discourse}}

@@

Problem-construction at the level of discourse is also more varied. They
are not monolithic in time, across location, or even by the same
speaker. While governmental discourses in Brazil have been studied for
topic such as inflation or race relations , we only find one systematic
analysis of Amazonian discourse. Barros (2020) investigates Amazonian
discourse in the Brazilian Congress with the objective of identifying
the arguments put forth by congressmen. The main finding is that in
congress the economic value of the Amazon for the cattle industry is the
most salient narrative, leading the author to conclude there is a
mismatch between the international debate (which focuses on
preservation) and the national debate (which focuses on economic
development).

While we find no major studies systematically looking into Amazonian
related themes in Brazilian presidential speeches, we find, for example,
several analyzes of environmental discourse in American presidential
speeches. Calderwood (2019) examines 2919 mentions of climate change in
American official presidential speeches since 1989. Among various
findings, one that stands out is that American presidents frequently
side-step the environmental aspects of climate change (ibid). He also
identifies a shift from economic to security framing of climate issues,
side-lining its environmental aspects. In another article, Calderwood
(2020) tests the effect of geographic location and type of communication
regarding climate change. Building prominently on Putnam (1988) but also
others, he hypothesizes that presidents are more likely to mention
climate change in foreign locations, and that location influences the
specific discursive approach and tone they adopt. He finds evidence in
support of his hypothesis, suggesting American presidential discourse at
the top on climate change does change based on location. Another example
is Bevitori (2015), who investigates how the `environment' has been
constructed in American presidential discourse since 1960 using a more
automated approach. The author finds that mentions of the environment
are typically co-selected with the pronoun `our', as well as with
`economy', `clean', and `preserve'.

@@

\hypertarget{theoretical-framework-problem-construction-and-presidential-discourse}{%
\subsection{2.2 Theoretical framework: problem-construction and
presidential
discourse}\label{theoretical-framework-problem-construction-and-presidential-discourse}}

In ``Journey towards Progress'', Hirschman (1963) analyzes three policy
problems in three different Latin American countries. The author draws a
conceptual distinction between pressing problems (pressured from outside
parties to the government) and chosen problems (chosen by the government
at their own discretion). Pressing problems can be either privileged or
neglected depending on the degree of pressure exercised by the
interested group. Problems can change from pressing to chosen across
time and in space as a function of solutions becoming available,
changing level of government control in society, or top policymakers
shifting interests(Hirschman 1963, 388--91). Choosing a problem, though,
entails a decision on how to represent it(Bacchi 2009). As Bacchi puts
``policy has a cultural dimension. It takes shape within specific
historical and national or international contexts'' (2009, 10). The
existence or proposal of a policy generally implies that there is a
(public) problem that needs (governmental) action to be fixed (ibid).
The alleged problem is not always explicitly stated in policy. Hirschman
exemplifies chosen problems with the case of the construction of
Brasilia (1975, 388). But building Brasilia can solve a problem of
regional inequality, a problem of a dormant economy without state
investment, a problem of political representation, or all three. How to
represent a policy is a matter of choice. And different representations
speak to different audiences.

@@

\hypertarget{research-design}{%
\section{3 Research Design}\label{research-design}}

\hypertarget{data-and-modeling-operationalizing-amazonian-problem-construction}{%
\subsection{3.1 Data and modeling: operationalizing Amazonian problem
construction}\label{data-and-modeling-operationalizing-amazonian-problem-construction}}

\hypertarget{analysis-and-limitations}{%
\subsection{3.2 Analysis and
limitations}\label{analysis-and-limitations}}

\hypertarget{how-has-the-amazon-been-constructed-as-a-problem}{%
\section{4 How has the amazon been constructed as a
problem?}\label{how-has-the-amazon-been-constructed-as-a-problem}}

@@

\hypertarget{the-rises-and-falls-of-the-amazon-as-a-topic-in-presidential-speeches}{%
\subsection{4.1 The rises and falls of the Amazon as a topic in
presidential
speeches}\label{the-rises-and-falls-of-the-amazon-as-a-topic-in-presidential-speeches}}

@@

\hypertarget{amazonian-problem-construction-in-time}{%
\subsection{4.2 Amazonian problem-construction in
time}\label{amazonian-problem-construction-in-time}}
